Corotating Dust Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-06-29 paper/ms.tex
update ms

\newcommand{\ntwoyear}{21} % TODO: update
\newcommand{\ntwoyearturnedoff}{5} % TODO: update

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust clumps in orbit at the
  Keplerian corotation radius.  Here, we report new CQVs discovered in
  TESS data collected between July~2018 and Sep~2022, and analyze
  their variability.  Our analysis of \nstarssearched\ K- and
  M-dwarfs with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\
  gold-standard CQVs.  Most of these discoveries are new, and they
  include the brightest and closest examples of this class of object
  known ($T$$\approx$11; $d$$\approx$20\,pc), as well as the most
  massive ($\approx$0.65\,M$_\odot$).  A few objects are outliers
  among outliers; LP 12-502 for instance shows a ``dip complex'' with
  a period and total duration that are fixed over more than 1000
  cycles, but which in detail shows anywhere from four to eight
  local minima per cycle.  LP-502 also demonstrated drastic shape
  changes over less than one cycle, and at times displayed distinct
  superposed periods simultaneously.  Broadly speaking, we find that
  none of the CQVs maintain a fixed light curve shape over timescales
  of more than a few hundred cycles -- they are all {\it
  quasi}periodic.  We revisit the arguments for why transient
  corotating dust clumps are the most likely explanation, though there
  are other possibilities.  In the future, we expect that our sample
  will help facilitate modeling and observational efforts aimed at
  understanding these objects, and connecting them to the broader
  contexts of star, disk, and exoplanet evolution.
\end{abstract}

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views
of \ntwoyear\ CQVs for which observations were available at least two years
apart.
Of the \ntwoyear\ CQVs for which such a baseline was available,
\ntwoyearturnedoff\ showed obvious signs of the phenomenon in one sector,
and marginal or non-existence signs in the other.
We emphasize that the periods themselves were all stable to $<$0.1\%.
Broadly speaking, this shape evolution
suggests that CQVs have a duty cycle of $\approx$75\%.

>>>>

\newcommand{\ntwoyear}{27} % TODO: update
\newcommand{\ntwoyearturnedoff}{5} % TODO: update

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust clumps in orbit at the
  Keplerian corotation radius.  Here, we report new CQVs discovered in
  TESS data collected between July~2018 and Sep~2022, and analyze
  their variability.  Our analysis of \nstarssearched\ K- and M-dwarfs
  with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\ gold-standard CQVs.
  Most of these discoveries are new, and they include the brightest
  and closest examples of this class of object known ($T$$\approx$9.5;
  $d$$\approx$20\,pc), as well as the most massive
  ($\approx$0.65\,M$_\odot$).  A few objects are outliers among
  outliers; LP 12-502 for instance shows a ``dip complex'' with a
  period and total duration that are fixed over more than 1000 cycles,
  but which in detail shows anywhere from four to eight local minima
  per cycle.  LP-502 at times demonstrated drastic shape changes over
  less than one cycle, and also displayed distinct superposed periods
  simultaneously.  Broadly speaking, we find that none of the CQVs
  maintain a fixed light curve shape over timescales of more than a
  few hundred cycles, and we revisit the arguments for why transient
  corotating dust clumps are the most likely explanation.  In the
  future, we expect that our sample will facilitate modeling and
  observational efforts aimed at understanding these objects, and
  connecting them to the broader contexts of star, disk, and exoplanet
  evolution.  For instance, based on the rate of dip evolution, we
  estimate that the incident dust flux corresponds to of order an
  asteroid belt's worth of mass accreting onto these stars over $10^8$
  years.
\end{abstract}

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CQVs
for which TESS 120-second cadence observations were available at least
two years apart.
Such a baseline was available for 32 of the \ngoods\ CQVs in our
catalog; for plotted purposes we show the brightest 27.
Of these 27 CQVs,
a few show clear signs of the phenomenon in one sector,
and marginal or non-existence signs in the other.
While there is some subjectivity in this assessment, to our eye
cases for which at least one sector would be flagged as ``ambiguous'' include
TIC~368129164 (S23 might be labelled an EB),
TIC~193831684 (which has been erroneously labelled an EB CITE Albrecht paper),
TIC~177309964 (S38 would be simply a rotating star),
TIC~404144841 (S38 looks like a rotating star),
TIC~201898222 (S3 looks like a rotating star),
TIC~144486786 (S32 might be flagged as a possible RS CVn),
TIC~38820496 (S28 might be a possible RS CVn).
One could easily also imagine in cases like TIC~206544316 that if
observed at lower signal to noise, the drastic shape evolution would
not be appreciated.
We emphasize that the periods themselves were all stable to $<$0.1\%.
Broadly speaking, this shape evolution
suggests that CQVs have a duty cycle of $\approx$75\%.
